---
title: 'Bayesian Stats HW3'
output:
  pdf_document: default
  html_document:
    df_print: paged
---

## Exercise4.1.
  In your write-up, include each line above and its results. Explain what each line does (in a bit more detail than the inline comments).
```{r}
# display the table class object
show(HairEyeColor) # Show data
# use apply function to sum the frequency by eye color and then by hair color
EyeHairFreq = apply( HairEyeColor, c("Eye","Hair"), sum ) # Sum across sex
# calculate the joint proportions by eye and by hair-color
EyeHairProp = EyeHairFreq / sum( EyeHairFreq )  # joint proportions, Table 4.1
# round the proportion to 2 decimal points and then display it
show( round( EyeHairProp , 2 ) )
# use apply function to sum the frequency by hair-color
HairFreq = apply( HairEyeColor , c("Hair") , sum ) # Sum across sex and eye
# calculate marginal proportions by hair-color
HairProp = HairFreq / sum( HairFreq ) # marginal proportions, Table 4.1
# round the proportion to 2 decimal points and then display it
show( round( HairProp , 2 ) )
# use apply function to sum the frequency by eye-color
EyeFreq = apply( HairEyeColor , c("Eye") , sum ) # Sum across sex and eye
# calculate marginal proportions by eye-color
EyeProp = EyeFreq / sum( EyeFreq ) # marginal proportions, Table 4.1
# round the proportion to 2 decimal points and then display it
show( round( EyeProp , 2 ) )
# calculate the conditional probability of hair color when eye color is blue 
EyeHairProp["Blue",] / EyeProp["Blue"] # conditional prob, Table 4.2
```
  Extend the above commands by also computing the probabilities of the hair colors given Brown eyes, and the probabilities of the eye colors given Brown hair.
```{r}
#the probabilities of the hair colors given Brown eyes
EyeHairProp["Brown",] / EyeProp["Brown"]
# the probabilities of the eye colors given Brown hair.
EyeHairProp[,"Brown"] / HairProp["Brown"]
```
## Ex 4.3
  A) There are 48/6 = 8 cards for value 10. The probability of getting a 10 is 8/48 = .167
  B) The probability of getting a 10 or a Jack is 8/48 * 2 = .333 because these two events are independent (mutually exclusive)
    
## Ex 4.6
```{r}
# marginal prob for grade 
PropGrade = c(.2, .2,  .6) # prob for 1st, 6th, and 11th grades respectively
PropGrade
# conditional prob table
condProp = data.frame("ice cream" = c(.3,.6,.3),
                      "fruit" = c(.6,.3,.1),
                      "french fries" = c(.1,.1,.6)) 
rownames(condProp) = c("p1st","p6th","p11th")
condProp
# joint prob is marginal prob * conditional prob
# p(food,grade) = p(food|grade) * p(grade)
jointprob = PropGrade*condProp
jointprob

```
  
  To see if grade and food are indepedent:
```{r}
# marginal prob for food
PropFood = colSums(jointprob)
PropFood
```
any cell in the joint probability table is not a product of their respective marginal probability. For instance,
$p(Ice cream) * p(1st grade) = .36\times0.2 =  0.072$  
which is not 0.06


