---
title: "Homework 6"
geometry: margin=1.2cm
output: pdf_document
---

```{r set up}
setwd("/Users//Bayesian Data Analysis/HW")
source("/Users/Bayesian Data Analysis/DBDA2Eprograms/BernGrid.R")
source("/Users/Bayesian Data Analysis/DBDA2Eprograms/DBDA2E-utilities.R")
source("/Users/Bayesian Data Analysis/DBDA2Eprograms/BernBeta.R")
source("/Users/Bayesian Data Analysis/DBDA2Eprograms/HDIofICDF.R")
```

## Class Example
The posterior distribution is more concentrated and narrow because the data distribution is more concentrated and we have more data which has more influence on posterior than prior
```{r, out.width ="50%"}
binwidth = 1/1000
thetagrid = seq(from = binwidth/2, to=1-binwidth/2, by= binwidth)
relprob = pmin(thetagrid, 1-thetagrid)
prior = relprob/sum(relprob)
plot(thetagrid, relprob)
plot(thetagrid, prior)

datavec = c(rep(1,200), rep(0,400-200)) #400 flips, 200 heads
posterior = BernGrid(Theta = thetagrid, pTheta = prior, Data = datavec)

```
  
## Exercises 6.1 
#### Part (A)
  The posterior is dbeta(theta|5,4) and is skewed towards head
```{r, out.width ="50%"}
posteriorA = BernBeta(c(4,4), c(1))
```
  
#### Part (B)
  The new posterior is dbeta(theta|6,4) and is skewed towards head
```{r, out.width ="50%"}
priorB = posteriorA
posteriorB = BernBeta(priorB, c(1))
```
  
#### Part (C)
  The new posterior is dbeta(theta|6,5) 
```{r, out.width ="50%"}
priorC = posteriorB
posteriorC = BernBeta(priorC, c(0))
```
  
#### Part (D)
  The new posterior is dbeta(theta|6,5) and is the same as before. The order doesn't matter.
```{r, out.width ="50%"}
# the same updates but in the order T, H, H
posteriorA.2 = BernBeta(c(4,4), c(0))
priorB.2 = posteriorA.2
posteriorB.2 = BernBeta(priorB.2, c(1))
priorC.2 = posteriorB.2
posteriorC.2 = BernBeta(priorC.2, c(1))
```
  
## Exercises 6.2
#### Part (A)
  The 95% HDI is .483 to .673.
```{r, out.width ="50%"}
prior.Ex2 = c(1,1)
data.Ex2 = c(rep(1,58),rep(0,100-58))
posterior.Ex2 = BernBeta(prior.Ex2, data.Ex2, showHDI = TRUE)
```
  
#### Part (B)
  The 95% HDI is .506 to .642 and is narrower than before.
```{r, out.width ="50%"}
prior.Ex2.2 = posterior.Ex2
data.Ex2.2 = c(rep(1,57),rep(0,100-57))
posterior.Ex2.2 = BernBeta(prior.Ex2.2, data.Ex2.2, showHDI = TRUE)

```





